Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-11-27 paper/main.tex
fixup! reorganize subsections; only mention heterogeneity in applications as paragraph
fixup! show data alongside heterogeneic growth rates for application 2

% --------------------------------------------------------------------------------------------------
\subsubsection{The role of Heterogeneity}

@@

However, we will now present data that indicate that an individual treatment is in fact necessary.
Figure~\ref{fig:estimated-growth-rates} shows the rod-lengths of individual bacteria which have been
calculated from the extracted positions.
We fit a delayed exponential function to them which is defined by

@@

where $\lambda$ is the growth rate and $t_0$ the delay.
Fitting this function to the individual growth curves of the bacteria and their mean growth curve,
both results in good agreements (Figure~\ref{fig:estimated-growth-rates} A,B).
However, the latter exhibits large amounts of uncertainty within the data which is also illustrated
in subfigure (C).
There, we can see much more clearly the difference between considering assigning parameters per cell
and global ones.
These results show that in order to accurately estimate the parameters of this system, we need to
account for this inherent heterogeneity and hence treat the growth parameters of the bacteria
individually.
\textbf{[Man ahnt was du meinst, doch leider ahnt man es nur. Vermeide Aussagen wie "it is clear".]}
\textit{Jetzt klarer?}

@@

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.28\textwidth]
        {figures/crm_fit-data-progression.png}
        \node at (0.02, 0.98)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.35\textwidth]
        {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-individual.pdf}%
        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.35\textwidth]
        {figures/crm_estimate_params/IWF-Goettingen/growth-rates-distribution.pdf}%
        \node at (0, 1)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}%
    \caption{
        \textbf{TODO Fix caption!}\\
        \textbf{Link to supplement for discussion of bar plot}\\
        % TODO use same y-axis across the plots
        (B) The subfigure shows the time-evolution of the rod-lengths for every individual
        bacterium.
        These growth curves are fitted individually with delayed exponential function.
        (C) The datapoints are created by averaging the rod-lengths shown in subfigure (A) and then
        fitted with an exponential function.
        The uncertainty band is the standard deviation of the averaged datapoints.
        (C) Blue dots show the combination of estimated growth rates and delays for individual
        bacteria while the orange dot corresponds to the average.
        Ellipses around the datapoints are taken from the uncertainty of the fits.
        Time is given in units of frames of the movie from which the data have been extracted.
    }
    \label{fig:estimated-growth-rates}
\end{figure}

>>>>

@@

However, we will now present data that indicate that an individual treatment is in fact necessary.
Figure~\ref{fig:application-2-data-overview} shows the rod-lengths of individual bacteria which have been
calculated from the extracted positions.
We fit a delayed exponential function to them which is defined by

@@

where $\lambda$ is the growth rate and $t_0$ the delay.
Fitting this function to the individual growth curves of the bacteria and their mean growth curve,
both results in good agreements (Figure~\ref{fig:application-2-data-overview} A,B).
However, the latter exhibits large amounts of uncertainty within the data which is also illustrated
in subfigure (C).
There, we can see much more clearly the difference between considering assigning parameters per cell
and global ones.
These results show that in order to accurately estimate the parameters of this system, we need to
account for this inherent heterogeneity and hence treat the growth parameters of the bacteria
individually.
\textbf{[Man ahnt was du meinst, doch leider ahnt man es nur. Vermeide Aussagen wie "it is clear".]}
\textit{Jetzt klarer?}

@@
